Sequential algorithms to split and merge ultra-high resolution 3D images

### 2017-07-31 paper.tex
Fixed (minor) error in multiple reads formulas. Explained them better.

Cluster reads are described in
Algorithm~\ref{algo:cluster-reads}. Function \texttt{switch} (line 3)
selects one of the three cases based on the amount of available memory
and the number of blocks. It returns \texttt{m'} and \texttt{case},
the identifier of the selected case. Function \texttt{check\_overlap}
(line 7) determines whether two blocks overlap multiple block slices
(case 2) or multiple block rows (case 1). For case 3 it always returns
true.  Function \texttt{sizeof} (line 8) returns the actual memory
used by its argument, including only its allocated segments in case
the argument is a buffer. It should be noted that writing the buffer
to the reconstructed image (line 11) might involve some manipulation
in the buffer to merge its continuous segments. \todo{Precise this
  based on Valerie's final code.}
\begin{algorithm}[h]
  \caption{Buffered merging of blocks with cluster reads}
  \label{algo:cluster-reads}
  \begin{algorithmic}[1]
    \STATE list = sort blocks by increasing (z,y,x) values
    \STATE initialize buffer
    \STATE (m',case)=\texttt{switch}(m,n,R,b)
    \STATE old\_block = list[0]
    \FOR{i = 0 ; i\textless n ; i+=1}
      \STATE block = list[i]
      \STATE overlap = \texttt{check\_overlap}(block,old\_block,case)
      \IF{\texttt{sizeof}(buffer)+\texttt{sizeof}(block) $\geq$ m' \textbf{or} overlap=true}
      \STATE write buffer in reconstructed image
      \STATE clear buffer
      \STATE overlap = false
      \ENDIF
      \STATE read block and insert it in buffer
      \ENDFOR
  \end{algorithmic}
\end{algorithm}

@@

In multiple reads, blocks are read partially to ensure that the memory
buffer only contains contiguous bytes, at any time. That is, a given
block will be read multiple times.     \todo{improve notations and algo in algorithm}

\begin{algorithm}[h]
  \caption{Buffered merging of blocks with multiple reads}
  \label{algo:multiple-reads}
  \begin{algorithmic}[1]
  \STATE sorted\_split\_list = sort splits based on its postion in increasing (z,y,x) values
  \STATE start\_index = 0, end\_index=(m-1)
  \STATE write\_range = (start\_index, end\_index)
  \WHILE{end\_index \textless sizeof(reconstructed\_image)}
    \STATE initialize data\_buffer
    \FOR{split in sorted\_split\_list}
      \IF{split is in write\_range}
        \STATE read split
        \STATE extract the rows in the write\_range
        \STATE write rows to the data\_buffer
      \ENDIF
    \ENDFOR
    \STATE write data\_buffer continuously to reconstructed\_image
    \STATE clear data\_buffer
    \STATE start\_index = end\_index + 1
    \STATE end\_index += mem
  \ENDWHILE

@@

For simplicity, we assume that $m'$ represents an integer number $k$
of block rows (case 1, $k<\sqrt[3]{n}$), of complete rows (case 2,
$k<d$), of tile rows (case 3, $k<\sqrt[3]{n}$), of slices (case 4,
$k<d$) or of block slices (case 5, $k<\sqrt[3]{n}$) \todo{make a
  figure}. In each of these 5 cases, we define $v$ as follows:
\begin{equation*}
  v_1=db \text{ ; }  v_2=Db \text{ ; } v_3=Ddb \text{ ; } v_4=D^2b \text{ ; } v_5=D^2db
\end{equation*}
so that we have:
\begin{equation*}
k_i=\floor*{\frac{m}{v_i}} \quad \mathrm{and} \quad m_i'=k_iv_i, \quad i \in \llbracket 1,5 \rrbracket
\end{equation*}
\todo{add one more preliminary step to explain how these equations were derived. Then jump to the final ones.}
Using these notations, the number of seeks performed by multiple reads is:
\begin{equation*}
  N_\mathrm{MR} =
\begin{cases}
  \floor*{\frac{Rb}{m_1'}}(k_1+1)+1
  + n \text{ mod } k_1& \text{if } d \leq \frac{m}{b} < D\\
  \ceil*{\frac{Rb}{m_2'}}(\sqrt[3]{n}+1)   & \text{if } D \leq \frac{m}{b} < Dd\\
  \floor*{\frac{Rb}{m_3'}}(k_3\sqrt[3]{n}+1)+1+\\
  \quad \quad \sqrt[3]{n}\left(\frac{n}{\sqrt[3]{n}} \text{ mod } k_3\right)&\
  \text{if } Dd \leq \frac{m}{b} < D^2\\
  \ceil*{\frac{Rb}{m_4'}}(\sqrt[3]{n}^2+1) & \text{if } D^2 \leq \frac{m}{b} < D^2d\\
  \floor*{\frac{Rb}{m_5'}}(k_5\sqrt[3]{n}^2+1)+1+&\\
  \quad \quad \sqrt[3]{n}^2\left(\frac{n}{\sqrt[3]{n}^2} \text{ mod } k_5\right)&\
  \text{if } D^2d \leq \frac{m}{b} < R
\end{cases}
\end{equation*}

which gives, using $R$ and $n$ as main variables:
\begin{equation}
N_\mathrm{MR} =
\begin{cases}
  \floor*{\frac{Rb}{m_1'}}\left(\frac{m_1'\sqrt[3]{n}}{\sqrt[3]{R}b}+1\right)+1+\\
  \quad \quad \quad n \text{ mod }\floor*{\frac{m\sqrt[3]{n}}{\sqrt[3]{R}b}}\
  & \text{if } \sqrt[3]{\frac{R}{n}} \leq \frac{m}{b} < \sqrt[3]{R}\\

@@

  \floor*{\frac{Rb}{m_3'}}\left(\frac{m'_3\sqrt[3]{n}^2}{\sqrt[3]{R}^2b}+1\right)+1+\\
  \sqrt[3]{n}\left( \frac{n}{\sqrt[3]{n}} \text{ mod } \floor*{\frac{m\sqrt[3]{n}}{\sqrt[3]{R}^2b}} \right)
  & \text{if } \frac{\sqrt[3]{R}^2}{\sqrt[3]{n}} \leq \frac{m}{b} < \sqrt[3]{R}^2\\

@@

  \ceil*{\frac{Rb}{m_5'}}\left(\frac{m_5'n}{Rb}+1\right)+1+\\
  \sqrt[3]{n}^2\left( \frac{n}{\sqrt[3]{n}^2} \text{ mod } \floor*{\frac{m\sqrt[3]{n}}{Rb}}\right)
  & \text{if } \frac{R}{\sqrt[3]{n}} \leq \frac{m}{b} < R
\end{cases} \label{eq:seeks-multiple-reads}
\end{equation}

>>>>

Cluster reads are described in
Algorithm~\ref{algo:cluster-reads}. Function \texttt{switch} (line 3)
selects one of the three cases based on the amount of available memory
and the number of blocks. It returns \texttt{m'} and \texttt{case},
the identifier of the selected case. Function \texttt{check\_overlap}
(line 7) determines whether two blocks overlap multiple block slices
(case 2) or multiple block rows (case 1). For case 3 it always returns
true.  Function \texttt{sizeof} (line 8) returns the actual memory
used by its argument, including only its allocated segments in case
the argument is a buffer. It should be noted that writing the buffer
to the reconstructed image (line 11) might involve some manipulation
in the buffer to merge its continuous segments. \todo{Precise this
  based on Valerie's final code.}
\begin{algorithm}[h]
  \caption{Buffered merging of blocks with cluster reads}
  \label{algo:cluster-reads}
  \begin{algorithmic}[1]
    \STATE sorted\_splits = sort splits by increasing (z,y,x) values
    \STATE initialize buffer
    \STATE (m',case)=\texttt{switch}(m,n,R,b)
    \STATE old\_block = list[0]
    \FOR{i = 0 ; i\textless n ; i+=1}
      \STATE block = list[i]
      \STATE overlap = \texttt{check\_overlap}(block,old\_block,case)
      \IF{\texttt{sizeof}(buffer)+\texttt{sizeof}(block) $\geq$ m' \textbf{or} overlap=true}
      \STATE write buffer in reconstructed image
      \STATE clear buffer
      \STATE overlap = false
      \ENDIF
      \STATE read block and insert it in buffer
      \ENDFOR
  \end{algorithmic}
\end{algorithm}

@@

Multiple reads are shown in Algorithm~\ref{algo:multiple-reads}.  The
main idea of this algorithm is that blocks are read partially (line 9)
to ensure that the memory buffer only contains contiguous bytes, at
any time. Therefore, the buffer can be written continuously to the
reconstructed image, without seeking (line 13).  It should be noted
that a given block might be read multiple times, in different memory
loads. 
\begin{algorithm}[h]
  \caption{Buffered merging of blocks with multiple reads}
  \label{algo:multiple-reads}
  \begin{algorithmic}[1]
  \STATE sorted\_splits = sort splits by increasing (z,y,x) values
  \STATE start\_index = 0 ; end\_index=(m-1)
  \STATE write\_range = (start\_index, end\_index)
  \WHILE{end\_index \textless Rb}
    \STATE initialize buffer
    \FOR{split in sorted\_splits}
      \IF{split has voxels in write\_range}
        \STATE split\_data = read split
        \STATE in split\_data, extract the rows in write\_range
        \STATE write rows to buffer
      \ENDIF
    \ENDFOR
    \STATE write buffer to reconstructed\_image
    \STATE start\_index = end\_index + 1
    \STATE end\_index += m
  \ENDWHILE

@@

For simplicity, we assume that $m'$ represents an integer number $k$
of block rows (case 1, $k<\sqrt[3]{n}$), of complete rows (case 2,
$k<d$), of tile rows (case 3, $k<\sqrt[3]{n}$), of slices (case 4,
$k<d$) or of block slices (case 5, $k<\sqrt[3]{n}$) \todo{make a
  figure}. In each of these 5 cases, we define $v$ as follows:
\begin{equation*}
  v_1=db \text{ ; }  v_2=Db \text{ ; } v_3=Ddb \text{ ; } v_4=D^2b \text{ ; } v_5=D^2db
\end{equation*}
so that we have:
\begin{equation*}
k_i=\floor*{\frac{m}{v_i}} \quad \mathrm{and} \quad m_i'=k_iv_i, \quad i \in \llbracket 1,5 \rrbracket
\end{equation*}

The total number of seeks performed by multiple reads in case $i$ is:
\begin{equation*}
  N^i_{\mathrm{MR}} = x_i + (x_i-1)b_i +b_i', \quad i \in \llbracket 1,5 \rrbracket
\end{equation*}
where $x_i$ is the total number of required memory loads, $b_i$ is the
number of blocks accessed by the first ($x_i$-1) memory loads and
$b_i'$ is the number of blocks access by the last memory load. We have:
\begin{equation*}
  x_i = \ceil*{\frac{Rb}{m_i'}} \quad ; \quad x_i -1 = \floor*{\frac{Rb}{m_i'}}, \quad i \in \llbracket 1,5 \rrbracket
\end{equation*}
and:
\begin{equation*}
b_1=k_1 ; \quad b_2=\sqrt[3]{n} ; \quad b_3 = k_3\sqrt[3]{n} ; \quad b_4=\sqrt[3]{n}^2 ; \quad b_5=k_5\sqrt[3]{n}^2
\end{equation*}
and:
\begin{eqnarray*}
b_1'=\sqrt[3]{n}D^2\text{ mod }k_1; &b_2'=b_2;   & b_3' = \sqrt[3]{n}\left( \sqrt[3]{n}D\text{ mod }k_3\right)\\
                                   &b_4'=b_4;  & b_5'=\sqrt[3]{n}^2\left( \sqrt[3]{n}\text{ mod }k_5\right) 
\end{eqnarray*}
It gives the following expression for $N_\mathrm{MR}$:
\begin{equation*}
  N_\mathrm{MR} =
\begin{cases}
  \floor*{\frac{Rb}{m_1'}}(k_1+1)+1
  + \sqrt[3]{n}D^2 \text{ mod } k_1& \text{if } d \leq \frac{m}{b} < D\\
  \ceil*{\frac{Rb}{m_2'}}(\sqrt[3]{n}+1)   & \text{if } D \leq \frac{m}{b} < Dd\\
  \floor*{\frac{Rb}{m_3'}}(k_3\sqrt[3]{n}+1)+1+\\
  \quad \quad \sqrt[3]{n}\left(\sqrt[3]{n}D\text{ mod } k_3\right)&\
  \text{if } Dd \leq \frac{m}{b} < D^2\\
  \ceil*{\frac{Rb}{m_4'}}(\sqrt[3]{n}^2+1) & \text{if } D^2 \leq \frac{m}{b} < D^2d\\
  \floor*{\frac{Rb}{m_5'}}(k_5\sqrt[3]{n}^2+1)+1+&\\
  \quad \quad \sqrt[3]{n}^2\left(\sqrt[3]{n} \text{ mod } k_5\right)&\
  \text{if } D^2d \leq \frac{m}{b} < R
\end{cases}
\end{equation*}
And finally, using $R$ and $n$ as main variables:
\begin{equation}
N_\mathrm{MR} =
\begin{cases}
  \floor*{\frac{Rb}{m_1'}}\left(\frac{m_1'\sqrt[3]{n}}{\sqrt[3]{R}b}+1\right)+1+\\
  \quad \quad \quad \sqrt[3]{n}\sqrt[3]{R}^2 \text{ mod }\floor*{\frac{m\sqrt[3]{n}}{\sqrt[3]{R}b}}\
  & \text{if } \sqrt[3]{\frac{R}{n}} \leq \frac{m}{b} < \sqrt[3]{R}\\

@@

  \floor*{\frac{Rb}{m_3'}}\left(\frac{m'_3\sqrt[3]{n}^2}{\sqrt[3]{R}^2b}+1\right)+1+\\
  \sqrt[3]{n}\left( \sqrt[3]{nR} \text{ mod } \floor*{\frac{m\sqrt[3]{n}}{\sqrt[3]{R}^2b}} \right)
  & \text{if } \frac{\sqrt[3]{R}^2}{\sqrt[3]{n}} \leq \frac{m}{b} < \sqrt[3]{R}^2\\

@@

  \ceil*{\frac{Rb}{m_5'}}\left(\frac{m_5'n}{Rb}+1\right)+1+\\
  \sqrt[3]{n}^2\left( \sqrt[3]{n} \text{ mod } \floor*{\frac{m\sqrt[3]{n}}{Rb}}\right)
  & \text{if } \frac{R}{\sqrt[3]{n}} \leq \frac{m}{b} < R
\end{cases} \label{eq:seeks-multiple-reads}
\end{equation}
